---
title: |
   | Business Analytics:
   | Sample Lecture 
subtitle: "2025 Open Day"
author: "Lachlan Deer"
institute: "University of Melbourne"
format: 
  beamer: 
    aspectratio: 32
    navigation: horizontal
    theme: mousteau
---

## The Business Analytics Process

\vspace{0.75cm}

> “Data are widely available; what is scarce is the ability to extract wisdom from them.” — Hal Varian, Chief Economist at Google

1.  **Defining the Problem** — What decision needs to be made? What are the options? What does success look like?

2.  **Collecting and Cleaning Data** — Gathering relevant data sources and ensuring they’re accurate, complete, and reliable.

3.  **Analyzing Data Appropriately** — Identify patterns, test hypotheses, and quantify trade-offs.

4.  **Communicating Insights** — Tell persuasive, relatable stories with data that clarify the decision at hand and guide action.

5.  **Acting to Create Value** — The goal is not analysis for its own sake, but to generate insights that drive strategy, improve operations, or better serve customers.


## Three Types of Business Analytics

**Fundamental goal:** to make smarter, more confident choices in an uncertain world.

-   **\alert{Different business problems require}** different types of insight—and therefore, **\alert{different analytical tools}**.

Broadly categorized into 3 types of business analytics

1.  Descriptive analytics
2.  Predictive Analytics
3.  **\alert{Causal Analytics}** -- the focus of this lecture

## 

```{r, echo = FALSE, fig.align = "center", out.width="80%"}
url <- "https://mailmeteor.com/assets/img/meta/email-personalization.png"
knitr::include_graphics(url)
```

## Lachlan's Inbox ...

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/lachlans_inbox.png"
knitr::include_graphics(url)
```

## Motivation

\alert{Business question}: Does adding personalized content improve email effectiveness?

**Why study this?** Personalized content $\rightarrow$ noninformative content

-   **Prevailing wisdom** is that noninformative content **should not influence consumer behaviour**
-   Most businesses engage in this practice
-   Email marketers have access to this information

## One way forward

How we're going to answer this question:

-   Run an **experiment in the field**
-   **Varies personalization** & across contexts

Discussion based on [**\alertc{Sahni, Wheeler and Chintagunta (2018)}**](https://pubsonline.informs.org/doi/10.1287/mksc.2017.1066)

## Field Experiments in One Slide

**Core idea of experiments**: \alert{Randomization into treatment}

```{r, echo = FALSE, fig.align = "center", out.width="45%"}
url <- "figs/random_allocation.png"
knitr::include_graphics(url)
```

PLUS the **measurement of outcome variable(s)**

-   Which have been (carefully) selected

**Comparison of average behaviour** between treatments **delivers causal effects**

-   Subject to some assumptions

**\alertc{ What about the "Field"}**? Run in **"real world" environments**, rather than computer labs

## Outcome Variable Selection

\vspace{-5cm}

\begin{center}
\textbf{\alert{What Outcomes Could We Measure?}}
\end{center}

## What Outcomes Could We Measure?

Email Marketers **have access to**:

-   Opens
-   Leads
-   Unsubscribes
-   Number of emails sent

And **can construct ratios** using these...

-   Opens / Sent
-   Leads / Opens

and so on ...

## Experiment 1: Context

```{r, echo = FALSE, fig.align = "center", out.width="60%"}
url <- "figs/experiment_1_email.png"
knitr::include_graphics(url)
```

\begin{center}
$\implies$ \textbf{Focus: acquiring new customers}
\end{center}

## Experiment 1: Design

**Control Group**: Typical email, no name added to subject line

Eg: "Learn Financial Modeling from Industry Experts"

**Treatment Group**: Names explicitly added to email's subject line

Eg: "Jack, Learn Financial Modeling from Industry Experts"

## Experiment 1: Results

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/experiment_1_results.png"
knitr::include_graphics(url)
```

## Mechanisms: Answering the "Why"

\vspace{-4cm}

**Lurking question**: Why did we observe this difference?

-   Academic wisdom suggests we shouldn't have seen differences
-   We **want to know why** a difference occurs
-   The "why" is known as \alert{"the mechanism"}

## Potential Mechanisms In the Email Context

The authors propose 3 mechanisms could be driving their results:

-   \[**A**\] Drawing **a**ttention to the email
-   \[**O**\] Shape **o**pinions of reciever / positive cue
-   \[**E**\] Increased **e**laboration via recipient processing

**Question 1**: Are these **really** the only possible mechanisms?

**Question 2**: What are the **consequences** of not finding the right mechanism?

## Mechanisms Predict Different Outcomes

Different mechanisms $\rightarrow$ different patterns in outcome variables

-   Increased **opens** of personalized emails
    -   Consistent with A, O, E
-   **Unsubscriptions**
    -   Increase under A, decrease under O and E
-   **Leads**
    -   increase under E and O ...
    -   ... more evidence for E if even more leads when a discount mentioned

## An Experiment to Isolate the Mechanism

```{r, echo = FALSE, fig.align = "center", out.width="70%"}
url <- "figs/experiment_4_design.png"
knitr::include_graphics(url)
```

**Doesn't mention discount size** due to worries of being spam filtered

## An Experiment to Isolate the Mechanism: Context

```{r, echo = FALSE, fig.align = "center", out.width="100%"}
url <- "figs/experiment_4_email.png"
knitr::include_graphics(url)
```

## An Experiment to Isolate the Mechanism:  Results

```{r, echo = FALSE, fig.align = "center", out.width="60%"}
url <- "figs/experiment_4_teststats.png"
knitr::include_graphics(url)
```

\vspace{-0.5cm}

**Statistical Tests**:

-   **A vs B**: adding name alone has no significant effect
-   **B vs D**: adding name in presence of discount increases leads

$\implies$ name increases engagement $\implies$ via elaboration mechanism

## Is it THAT easy? ... Likely Not

```{r, echo = FALSE, fig.align = "center", out.width="60%"}
url <- "figs/marketing_letters_title.png"
knitr::include_graphics(url)
```

```{r, echo = FALSE, fig.align = "center", out.width="60%"}
url <- "figs/marketing_letters_abstract.png"
knitr::include_graphics(url)
```

Paper: [**\alertc{Defau \& Zauner (2023)}**](https://link.springer.com/article/10.1007/s11002-023-09701-7)

## Is it THAT easy? ... Likely Not

\begin{center}
Marketers re-'discovered' \textbf{\alert{The Novelty Effect}}
\end{center}

"...But in the last years, this \alertc{marketing trick has become a common practice} in email marketing, and therefore, it might be that the behavior of recipients changed as they got used to it... \alertb{we find no indication} that using the **first name of the recipient** in the subject line **has a positive effect** on opening rates orclick-through rates"

BUT

"... we do find that **using the surname of the recipient has a positive effect on opening rates and click-through rates**; an indication that similar marketing techniques still have potential, but that they need to be more creative and surprising"


## Want to learn more about the BA program?


\begin{center}
  \includegraphics[width=0.45\linewidth]{figs/qr_code_ba.png}

  \textbf{The Business Analytics Major}
\end{center}

## Want to learn more about this class?

\begin{center}
  \includegraphics[width=0.45\linewidth]{figs/qr_code_fba.png}

  \textbf{CMCE-10002: Foundations of Business Analytics}
\end{center}
